\documentclass[11pt]{book}

\begin{document}

\chapter{Modules and How They Work}

Modular synthesis is an exciting, freeing way to create music and design sounds.

All Audulus modules are created using Audulus nodes. You can open up any module to see how it works. Some modules are very simple and may have only a few nodes inside. Others are a web of interconnected submodules, sometimes many layers deep.

The great thing about the Audulus module library is that you don't need to understand how each module is constructed to be able to use it. There are, however, a few things you need to know about the module library before getting started.

\section{Signal Standards}

Modules in the Audulus library operate with a set of standardized signals. The job of the modules is to generate and modify these signals in interesting ways.

There are four main signal types: Gate, Modulation, 1/Octave, and Audio. As a general rule, you will connect like with like, meaning connect a Gate output to a Gate input, or an Audio output to an Audio input.

\subsection{Gate}

The Gate signal is used to drive sequencers, open and close envelopes, and syncronize tempo, among other things. Its signal range is 0 or 1: off or on.

Gates are generated mainly by clock modules and keyboard or MIDI input. You can modify gate signals with modules like clock dividers and multipliers, gate sequencers, and Bernoulli gates. They are used by modules like sequencers, envelopes, and LFOs.

Gate inputs and outputs are always marked with a green light which is also called the Light node. When a gate input or output is equal to 0, the color of the light is dark green, and when the output is equal to 1, the color of the light is a bright green.

The only exception to this rule is when a gate is entering an envelope. Here, gate height, or the upper value of the gate, will set the maximum value that the Attack period rises to. This allows you to play with dynamics, or making your sounds softer or louder.

Gate inputs and outputs may be marked Gate, Gte, Reset or Rset, PWM, or Clk. Sometimes they may be unmarked except for the light node inside each input or output. These labels will have contextual meanings for each specific module, which are explained in their individual entries in the manual below.

Most modules will respond to only the rising edge of a gate. The rising edge is the moment where a gate transitions from 0 to 1. For example: a step sequencer will only step forward at the rising edge, but will not step forward again on the falling edge.

Other modules, like an envelope, respond to both the rising and falling edge of the gate signal. The rising edge of the gate will initiate the attack stage of the envelope whereas the falling edge will initiate the release stage.

If you are familiar with modular synthesis, you might be wondering if there is a difference between a pulse and a gate in Audulus. A pulse is a very short on/off burst, usually generated by clock modules, whereas gate signals can be any length from very short to very long.

In Audulus, no distinction is made between a pulse and a gate. That said, several clock modules give you control over the pulse width of their gate output. Pulse width is the ratio of on to off time. A 10\% pulse-width gate is on 10\% of the time and off 90\% of the time whereas a 90\% pulse-width gate is on 90\% of the time and off 10\% of the time.

\subsection{Modulation}

The Modulation signal is used to tweak parameters like filter cutoff, VCO shape, and VCA level. Its signal range is 0 to 1. 

Modulation signals are genereated by modules like LFOs, envelopes, sequencers, and sample \& holds. They can be used to modulate any knob on any module and often have their own separate inputs as well.

Modulation inputs and outputs are often marked with a red light. The light indicates the relative strength of the incoming modulation. If the light is black (not lit) its value is 0. If the light is fully red its value is 1.

When a module has several outputs like the Basic LFO, lights may be omitted to save CPU time. If a portion of a module does not terminate in a meter or audio output, then everything that precedes it will not be calculated. This means if you use only the Sine output of the Basic LFO, only the sine path will be calculated. If lights were present at each output they would force Audulus to calculate the unused paths, wasting CPU time.

Modulation inputs and outputs may be marked Mod, Env (Envelope), a specific waveshape like Sine or Saw, or correspond to a knob label like Hz Mod or Q. Sometimes they may be unmarked and have contextual meanings for each specific module, which are explained in their individual entries in the manual below.

\subsection{1/Octave}
\subsection{Audio}

\section{Categories of Modules}

\section{Wiring Modules Together}
\subsection{Sequencer-Driven Subtractive Synthesizer}
\subsection{Keyboard-Controlled FM Synthesizer}
\subsection{Drum Patterns and Mixing}
\subsection{Stereo Effects}
\subsection{Generative Sequencing}
\subsection{Automation}
\subsection{DAW Integration}
\subsection{Eurorack Integration}


\chapter{Clock Modules}
\section{Clock Divider}
\section{Master Clock}
\section{Bernoulli Gate}
\section{Chance Select Gate}
\section{Pachinko Machine}
\section{Random Bursts}
\section{uClock}

\chapter{Drum Modules}
\section{uKick}

\chapter{Effect Modules}
\section{Delay}
\subsection{Stereo Delay}
\subsection{Looper Sync}
\subsection{Tempo Sync}
\subsection{uDelay}
\section{Distortion}
\subsection{Asymmetrical Drive}
\subsection{Fold Processor}
\section{Dynamics}
\section{Reverb}
\subsection{Really Humungous Reverb}
\subsection{uVerb}

\chapter{Envelope Modules}
\section{Analog Envelope}
\section{EOC Max ADSR}
\section{uADSR}

\chapter{Input-Output Modules}
\section{Audio Input}
\section{Audio Output}
\section{Expert Sleepers ES-3}
\section{Expert Sleepers ES-6}
\section{Expert Sleepers ES-8}
\section{MIDI Input}
\section{VPO Converter}

\chapter{LFO Modules}
\section{Basic LFO}
\section{Octature Sine LFO}
\section{TZFM LFO}
\section{uLFO}

\chapter{Mixer Modules}
\section{uMix Audio}
\section{uMix CV}

\chapter{Quantizer Modules}
\section{Chromatic Quantizer}
\section{Gateable Quantizer}
\section{Modulation Quantizer}
\section{Neo-Reimannian Triad Transformer}
\section{Quick Quantizer}

\chapter{Sequencer Modules}
\section{8 Step Sequencer}
\section{16 Step Gate Sequencer}
\section{Binary Gate Sequencer}
\section{Euclidean Gate Sequencer}
\section{Matrix Gate Sequencer}
\section{Shape Gate Sequencer}

\chapter{Utility Modules}
\section{Attenuate-Offset}
\section{Attenuverter}
\section{Audio Attenuverter-Offset}
\section{Automation Lane}
\section{Chaos Generator}
\section{Modulation to 1/Oct}
\section{Sample \& Hold}
\section{Slew}
\section{Switch Sequential 8 Step}
\section{VC Switch}
\section{uLogic}
\section{uSample \& Hold}
\section{Window Comparator}

\chapter{VCA Modules}
\section{Digital VCA}
\section{Diode VCA}
\section{uVCA}

\chapter{VCF Modules}
\section{Diode Ladder VCF}

\chapter{VCO Modules}
\section{Basic VCO}
\section{Chebyshev Additive VCO}
\section{Karplus-Strong VCO}
\section{Morphing VCO}
\section{Noise}
\section{Skew VCO}
\section{Supersaw VCO}
\section{TZFM VCO}
\section{uNoise}
\section{uVCO}







\end{document}
